PhD Title \\ PhD Thesis

### 2024-08-02 Latex_files/Thesis/chapters/chapter 3/main.tex
tidying up chapter 3

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the PhD was to investigate 
the possibility of using optical tweezing as a means of initiating and 
controlling nucleation by generating fluid flow within a small droplet of 
supersaturated solution. The goal of which would be to understand the 
influence of shearing on nucleation at a micro level as compared to larger 
scale results. It has been shown that for macro-scale systems, the likelihood 
of nucleation increases to a maximum value under increased shearing 
\cite{Debuysschere2023, Mura2016}. Theoretical research into the matter 
identified two competing effects that effect a crystal in moving fluid fields; 
firstly, nucleation is enhanced due to the increased mass transfer of solute 
material; and secondly, shear flow against the crystal surface leading to a 
decrease in growth \cite{Mura2016}. These two competing effects are validated 
by experimental work using glycine solutions, showing that beyond a certain 
shear rate the nucleation rate is reduced \cite{Debuysschere2023}. In this 
chapter I outline the optical tweezer equipment used during this PhD, the 
initial attempts made to induce nucleation via optical rotation, and the 
impact of a moving beam on the crystal growth of a newly formed nucleus.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, two
microscope objectives (one for trapping and one for imaging),
and a means of controlling the position of the loaded sample. 
While there are other pieces of equipment used in modern tweezer
experiments these are the bare minimum requirements for any 
optical tweezer. The laser used for this project was a 1064 
nm near infrared laser - provided by CNI Lasers – that has 
an adjustable power supply to vary the energy output of the 
laser - the remaining optics where supplied by Thor Labs.
Experimental work has shown that the trapping efficiency 
increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ 
\cite{kim2003dependence} where $D_{obj}$ is the diameter
of the objective aperture. To expand the beam front we 
utilise a Galilean beam expansion arrangement (indicated 
by $f_1$, and $f_2$) as recommended for high power laser 
applications. In our initial experiments the beam expansion 
provides a $4\times$ magnification; however, in later experiments 
where we utilise a galvano-mirror the beam expansion is $3\times$ 
and then the 4f correlator provides a further $1.25\times$ magnification 
(using $f_3$ and $f_4$) - where the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

It should be noted that the galvano-mirror requires the use 
of a Keplerian beam expansion arrangement which reduces
the transmitted laser power due to localised heating of the air.
Afterwards the laser is passed through a dichroic mirror that 
separates incoming infrared and visible light, this is to 
prevent the laser from damaging the CCD camera used for imaging
the trapping plane. The laser is then focused to a diffraction
limited spot by a 1.25 NA objective. By increasing the numerical
aperture of the objective, the gradient force at the focal point 
is increased; the trade-off being that the for higher NA values 
the trapping depth is reduced due to spherical aberrations.
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses
the scattered laser light and also provide an aperture for an 
imaging LED to illuminate the focal plane. Samples are loaded 
onto a piezo driven table to that is inserted between the trapping
and condensing objectives; the piezo drivers allow for sub-micron
control of the sample position to a degree as small as a 10 nm. 
To detect and monitor the position of a trapped particle a 
quadrant photo diode (QPD) was utilised. 
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively. Diagram not drawn to scale.}
	\label{fig:setup}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, 
a position detection system is utilised. There are 3 possible
methods of position detection: video-analysis, lateral-effect
position sensing, and photodiodes. The former being 
ideally suited for multiple traps or situations where precision
is not the top priority. In order to match the force measurements
of back-focal plane interferometry requires the camera's frame
rate to exceed $1\ kHz$ which can be difficult to achieve while 
maintaining a decent resolution \cite{Gibson2008}. In comparison
off the shelf back-focal plane detectors can achieve temporal
resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.

A QPD is frequently used position detection system for optical 
tweezers due to their high sampling rate, high degree of 
precision, and ease of set up. The QPD is constructed of four photo 
diodes assembled in a quadrant formation, when a particle is trapped the 
interference pattern produced is focused onto the QPD, with 
the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants 
together the particle's centre of mass is tracked in the x-y 
plane. Axial displacement can be estimated by observing the change
in the total voltage of the QPD. The outputted signal gives an
indication of the particle's relative displacement from the beam 
focus, but in order to convert the signal to distance units the
trap needs to be calibrated (assuming a linear response curve).
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

A lateral-effect sensor has a similar output but works using a 
the entire sensor as a single cell analogous to the focal plane of
the trapping beam. The four corners of the sensor act as anodes 
connected to a base plate cathode, as the beam moves across the
surface of the detector each anode will experience a different 
photocurrent depending on how close the centre of the 
interference pattern is to each anode. The advantage of a 
lateral effect detector is that the linear regime is much larger 
than a QPD making it much better for monitoring the position of a
trapped particle. However, Lateral-effect sensors are often limited
in their spacial resolution due to high signal-to-noise ratios, 
requiring a high intensity of light on the sensor in order to 
get a clean signal. As a result, most optical trapping experiments
are conducted using a QPD as opposed to a lateral-effect sensor, 
as often the displacement is small enough that the QPD response 
curve can be considered linear.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
Generation of fluid shear can be achieved via two avenues: Firstly, by 
utilising circularly polarised light it is possible to transfer angular 
momentum from the laser to the trapped entity. Secondly, one can 
directly move the trap within the imaging plane by steering the beam 
using either a galvanometric mirror or gimbal mirror. The following 
chapter outlines the work done with shear generated by circularly 
polarised light and the challenges of applying this to localising 
nucleation.

There are several options for particles that can be rotated using 
optical tweezers \cite{Parkin2009, Saito2022}. Over the course of 
the PhD two different micro spheres where investigated, vaterite 
and liquid crystal droplets. Both can be readily synthesised in 
the lab and are will rotate at a variety of sizes. While silver 
nano particles were considered their high cost and small size 
meant they were disregarded as an option for optical rotation.

Vaterite is a polymorph of calcium carbonate that is rarely seen in 
nature due to its low stability. However unlike its other polymorphs of 
calcite and aragonite, when synthesised vaterite will typically form 
small spherical particles making them ideal for optical trapping and rotation. 
Synthesis of vaterite micro spheres requires fine control of the nucleation
process in order to maintain polymorphic stability, though for the purposes
of optical rotation the exact polymorph is not as important as its morphology
as all 3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts of 
$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of Agepon was added
to halt the reaction, the solution was filtered and washed 3 times with 
distilled water before being suspended in water. When trapped in circularly 
polarised light, the anisotropic scattering of the sphere results in a periodic 
signal on the QPD. Therefore, the resulting power spectrum is not a Lorentzian
but now also displays peaks that appear at integer multiples of the particles
rotational frequency. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
	\label{fig:vaterite}
\end{figure}

As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
demonstrates a Lorentzian curve but modified with these periodic peaks, 
while the Lorentzian can be loosely fitted to the end tail there exists 
no current model for describing the power spectra. The closest approximation
to this was conducted by \cite{Yogesha2012} where they describe the 
rotational motion of ellipsoidal polystyrene particles. The critical assumption
being that the particle perfectly rotates in the $x-y$ plane; however it
has long been suspected that birefringent microspheres experience random
rotations outside of the $x-y$ plane \cite{Bang2020} making it very difficult 
to characterise the forces on rotating microspheres without a proper understanding 
of the optical torque being applied to it.

\subsection{Liquid Crystal Rotors}
Liquid crystals are an intriguing example of materials with mixed phase
properties, unlike typical solutes such as Glycine, a liquid crystal
can still maintain some degree of order between its individual molecules
while in the liquid state. This is due to the fact that liquid crystals
are constructed of long ordered molecules that demonstrate a long range
ordering. There are three main types of liquid crystal transition methods:
thermotropic crystals will transition to their liquid crystal phase when 
sufficiently heated; Lyotropic materials can undergo this transition due 
to changes in temperature and concentration; and lastly Metallotropic 
materials - which are composed of both organic and inorganic molecules - 
change phase according to the ratio of organic to inorganic molecules present.
Liquid crystal rotors are rather simple in their production, 
4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
and a small amount was added to a vial of distilled water. The solution 
was then heated in a water bath to $25^\circ$ in order to transition the
solid crystal into its liquid crystal state. The individual droplets 
are inherently birefringent and rotate in a circularly polarised trap. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\linewidth]{LC_sample.png}
	\caption{Liquid crystal undergoing rotation due to the circularly polarised trap.} 
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
Using a birefringent particle and rotating it within the fluid, the maximum 
rotation rate is due to the fluid drag resisting the torque of the 
trapping beam \cite{RodriguezSevilla2018}. If you want to measure 
fluid flow you can instead use a micro-rotor to see how fluid flow 
propagates in the medium \cite{Knoener2005}. Likewise, one can use 
a galvanometric mirror to probe the drag force of the fluid, by 
understanding the trap strength (calibrating using a low frequency 
signal) one can measure the drag force experienced by the local fluid \cite{RobertsonAnderson2018}. I

Understanding the fluid velocity around our trapped object is determined 
mostly by the Reynold's number of our system, for a sphere submersed 
in a moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}

@@

Where $d$ is the particle diameter, $\omega$ is the rotation frequency
reported by the QPD, and $r$ is the distance from the particle's centre. 
Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating
outward from the centre of the sphere. The shear rate can then be computed
as the partial derivative fluid flow (assuming shearing is generated purely
by the flow field):
\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}

First we tested the rotational behaviour of vaterite in distilled water, 
samples of vaterite where diluted down and $200\mu L$ was pipetted onto
the sample stage. Due to Van der Waal's forces some of the microspheres 
were stuck together, fortunately individual sphere's were still present.
Over the course of a few weeks multiple sphere's where trapped 
and their rotation speeds were measured via the QPD. 
\begin{figure}[h!]
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a) Fluid flow radiating out from the surface of a rotating vaterite sphere. (b) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ is indicated by the dotted line. Vaterite radii and rotation frequencies are shown, the laser power was kept constant at 450 mW. Reported rotation rates, and their corresponding fluid flow and shear rates, for vaterite are also plotted alongside lab results.}
	\label{fig:vaterite_shear}
\end{figure}

From Fig.\ref{fig:vaterite_shear} there is not a strong relationship between
particle size and rotation rate, this is contrary to much of the theoretical
predictions that predict an exponential decay with particle size. This can
be in part due to the fact that synthesising perfectly spherical spheres that
have uniform birefringence is difficult over multiple experiments and tests. 
Despite our best efforts at controlling the growth rate the vaterite spheres
would often combine together while suspended in water after a short period of 
time. The fastest reported rotation rate found during this PhD was by \cite{Arita2016}
that achieved a rotation rate of $5\ kHz$, this is plotted on Fig~\ref{fig:vaterite_shear}
as the dotted line. Even at that extreme a rotation rate the region in which 
nucleation is at its optimal likelihood is only $20\ nm$ wide meaning that 
localising nucleation around a rotating sphere is vanishingly small.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of micro-rotors in supersaturated solutions}
Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
(see sec.~\ref{sec:vaterite}), and then suspended in distilled 
water, at the same time a supersaturated solution of Glycine 
and water was prepared and $50\ \mu L$ was pipetted onto a 
glass cover slip. A single microsphere the vaterite was trapped 
in a circularly polarised light and brought close to the droplet 
edge.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
	\caption{Diagram of optical trapping set up for rotating birefringent 
		particles in a supersaturated solution. Left: side view of the 
		trapping set up showing the location of the trap focus at the 
		edge of the droplet of a supersaturated solution. Right: top down 
		view of the glycine droplet with a trapped birefringent particle 
		shown close to edge of the trap. As the particle rotates the drag 
		force from the surrounding fluid generates a flow field around 
		itself (see Eq.~\ref{eq:birefringent_speed}).}
\end{figure}

>>>>

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the project is to 
investigate the possibility of using optical tweezing to induce nucleation 
by generating fluid flow within a supersaturated solution. The intent 
of which would be twofold: Firstly to have a repeatable means of 
inducing nucleation under different solution conditions. And secondly, 
to understand the influence of shearing on nucleation at a micro 
level as compared to results in bulk fluid. It has been shown that 
for macro-scale systems, the likelihood of nucleation increases to 
a maximum value under increased shearing \cite{Debuysschere2023, 
Mura2016}. Mura and Zaccone developed a theoretical framework to 
describe how the a newly formed nucleus experiences two additional 
growth factors when placed in a moving fluid. Firstly, due to increased 
molecular transport of solute molecules the nucleation rate is enhanced 
in low to moderate fluid flows. But in addition, due to shear flow the 
crystal surface undergoes deformation which suppresses the nucleation 
rate undergoing faster fluid flow \cite{Mura2016}. The theory was 
confirmed for glycine solution by Debuysschere \textit{et al} who
identified that the nucleation rate of supersaturated glycine was 
enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ \cite{Debuysschere2023}.
After which the nucleation rate began to decrease but was 
still greater compared to the case where fluid flow was minimal.

Optical tweezers can been used to rotate a whole host of micro-rotors, 
with the fastest reported results exceeding $1000\ Hz$ in heavy water
\cite{Arita2016}. Therefore, it stands to reason that a micro-rotor
rotating in a supersaturated solution could generate sufficient fluid
flow such that the nucleation rate around said particle is enhanced 
compared to the bulk fluid. We focused on two primary candidates for
micro-rotors, Vaterite and 4-Heptyl-4-biphenylcarbonitrile (7CB). The
former being a polymorph of calcium carbonate and the latter an example
of nematic liquid crystals, both of which have been used repeatedly in
previous micro-rotor. In addition, we also consider the application of
using techniques beam steering to generate fluid flow by trapping silica
micro-beads. In this instance the fluid flow is generated not due to the
transfer of angular momentum, but due to shearing caused by a moving 
sphere through stagnant fluid.

To begin with, the discussion of the necessary optical equipment is 
covered, drawing attention to specialised components and techniques 
that are not standard in optical trapping set ups.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, a focusing 
microscope objectives, a position controller, and position detector.
The laser used for this project was a $1064\ nm$ near infrared laser 
- provided by CNI Lasers – that was focused by a Nikon 100x oil 
immersion lens. The choice of an oil immersion lens is important as 
the optical oil used prevents a loss of focus when used on a glass 
cover slip. Now, experimental work has shown that the trapping 
efficiency increases with beam diameter up until it exceeds 
$\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where $D_{obj}$ is 
the diameter of the objective aperture. To expand the beam front we 
utilise a Galilean beam expansion arrangement (indicated by $f_1$, 
and $f_2$ in Fig.~\ref{fig:setup}) as recommended for high power laser 
applications. In our initial experiments the beam expansion provides a 
$4\times$ magnification. Whereas in later experiments we utilised a 
galvano-mirror the beam expansion is $3\times$ and then the 4f 
correlator provides a further $1.25\times$ magnification (using $f_3$ 
and $f_4$) - the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

It should be noted that the galvano-mirror requires the use of a 
Keplerian beam expansion arrangement which reduces the transmitted 
laser power due to localised heating of the air. Afterwards the 
laser is passed through a dichroic mirror that separates incoming 
infrared and visible light, this is to prevent the laser from 
damaging the CCD camera used for imaging the trapping plane. The 
laser is then focused to a diffraction limited spot by the objective. 
Utilizing a high numerical aperture objective enhances the gradient 
force at the focal point; the trade-off being that the for higher NA 
objectives the trapping depth is reduced due to spherical aberrations. 
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses 
the scattered laser light and also provide an aperture for an imaging 
LED to illuminate the focal plane. Samples are loaded onto a piezo 
driven table to that is inserted between the trapping and condensing 
objectives; the piezo drivers allow for sub-micron control of the 
beam focus position to a degree as small as a 10 nm. To detect and 
monitor the position of a trapped particle a quadrant photo diode 
(QPD) was utilised. 
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively. Diagram not drawn to scale.}
	\label{fig:setup}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, a 
position detection system is required. There are 3 possible methods 
of position detection: video-analysis, lateral-effect position sensing, 
and photodiodes. The former being ideally suited for multiple traps 
or situations where precision is not the top priority. In order to 
match the force measurements of back-focal plane interferometry requires 
the camera's frame rate to exceed $1\ kHz$ which can be difficult 
to achieve while maintaining a decent resolution \cite{Gibson2008}. 
In comparison off the shelf back-focal plane detectors can achieve 
temporal resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.

A quadrant photo diode (QPD) is a frequently used position detection 
system for optical tweezers due to their high sampling rate, high 
degree of precision, and ease of set up. The QPD is constructed of 
four photo diodes assembled in a quadrant formation, when a particle 
is trapped the interference pattern produced is focused onto the QPD, 
with the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants together 
the particle's centre of mass is tracked in the x-y plane. Axial 
displacement can be estimated by observing the change in the total 
voltage of the QPD. The outputted signal gives an indication of the 
particle's relative displacement from the beam focus, but in order to 
convert the signal to distance units the trap needs to be calibrated 
(assuming a linear response curve).
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

A lateral-effect sensor has a similar output but works using a the 
entire sensor as a single cell analogous to the focal plane of the 
trapping beam. The four corners of the sensor act as anodes connected 
to a base plate cathode, as the beam moves across the surface of the 
detector each anode will experience a different photocurrent depending 
on how close the centre of the interference pattern is to each anode. 
The advantage of a lateral effect detector is that the linear regime 
is much larger than a QPD making it much better for monitoring the 
position of a trapped particle. However, Lateral-effect sensors are 
often limited in their spacial resolution due to high signal-to-noise 
ratios, requiring a high intensity of light on the sensor in order to 
get a clean signal. As a result, most optical force measurements are 
conducted using a QPD as opposed to a lateral-effect sensor, as often 
the displacement is small enough that the signal-displacement curve 
can be considered linear.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
There are several options for particles that can be rotated using optical 
tweezers \cite{Parkin2009, Saito2022}. Over the course of the project two 
different micro spheres where investigated, vaterite and liquid crystal 
droplets. Both can be readily synthesised in the lab and are will rotate 
at a variety of sizes.

Vaterite is a polymorph of calcium carbonate that is rarely seen in nature 
due to its low stability \cite{KonopackaLyskawa2019}. However unlike its 
other polymorphs of calcite and aragonite, when synthesised vaterite will 
typically form small spherical particles making them ideal for optical 
trapping and rotation. Synthesis of vaterite micro spheres requires fine 
control of the nucleation process in order to maintain polymorphic stability, 
though for the purposes of optical rotation the exact polymorph is not as 
important as its morphology as all 3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts of 
$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of Agepon was added
to halt the reaction, the solution was filtered and washed 3 times with 
distilled water before being suspended in water.

When trapped in circularly polarised light, the anisotropic crystal lattice
allows spin angular momentum to be transferred to the vaterite particle, 
resulting in a rotation about the beam axis. In addition, the anisotropic 
scattering causes the QPD signal to vary with a constant periodicity that
is attributed to its rotation rate. Therefore, the resulting power spectrum 
is not a Lorentzian but now also displays peaks that appear at integer 
multiples of the particles rotational frequency. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
	\label{fig:vaterite}
\end{figure}

As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
demonstrates a Lorentzian curve but modified with these periodic peaks, 
while the Lorentzian can be loosely fitted to the end tail there exists 
no current model for describing the power spectra. The closest approximation
to this was conducted by \cite{Yogesha2012} where they describe the 
rotational motion of ellipsoidal polystyrene particles. The critical 
assumption being that the particle perfectly rotates in the $x-y$ plane; 
however it has long been suspected that birefringent microspheres 
experience torques outside of the $x-y$ plane \cite{Volpe2023} making 
it very difficult to characterise the behaviour of rotating birefringent 
microspheres without a proper understanding of the full optical torque 
being applied to it.

\subsection{Liquid Crystal Rotors}
Liquid crystals are an intriguing example of materials with mixed phase
properties. Unlike typical solutes such as Glycine, a liquid crystal
can still maintain some degree of order between its individual molecules
while in the liquid state. This is due to the fact that liquid crystals
are constructed of ordered molecules that demonstrate a long range
ordering. There are three main types of liquid crystal transition methods:
Thermotropic crystals will transition to their liquid crystal phase when 
sufficiently heated. Lyotropic materials can undergo this transition due 
to changes in temperature and concentration. And lastly, Metallotropic 
materials - which are composed of both organic and inorganic molecules - 
change phase according to the ratio of organic to inorganic molecules present.
Liquid crystal rotors are rather simple in their production, 4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
and a small amount was added to a vial of distilled water. The solution 
was then heated in a water bath to $25^\circ$ in order to transition the
solid crystal into its liquid crystal state. The solution can then be 
loaded onto a sample cover slip and the individual droplets visualised.
The molecules of 7CB will align with a strong electric field, and due to
the spherical droplet geometry the droplets are inherently birefringent. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\linewidth]{LC_sample.png}
	\caption{Liquid crystal undergoing rotation due to the circularly polarised trap.} 
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
Using a birefringent particle and rotating it within the fluid, the maximum 
rotation rate is due to the fluid drag resisting the torque of the trapping 
beam \cite{RodriguezSevilla2018}. If you want to measure fluid flow you can 
instead use a micro-rotor to see how fluid flow propagates in the medium \cite{Knoener2005}. Likewise, one can use a beam steering arrangement to 
probe the drag force of the fluid, by understanding the trap strength 
(calibrating using a low frequency signal) one can measure the drag force 
experienced by the local fluid \cite{RobertsonAnderson2018}. I

Understanding the fluid velocity around our trapped object is determined 
mostly by the Reynold's number of the system, for a sphere submersed in a 
moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}

@@

Where $d$ is the particle diameter, $\omega$ is the rotation frequency
reported by the QPD, and $r$ is the distance from the particle's centre. 
Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating
outward from the centre of the sphere. The shear rate can then be computed
as the partial derivative fluid flow (assuming shearing is generated purely
by the flow field):
\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation of fluid flow around micro-rotors}
First we determined the upper rotation rate that could be achieved 
using both Vaterite and liquid crystal spheres. Vaterite samples were 
synthesised according to \cite{Parkin2009, Bishop2004} (see sec.~
\ref{sec:vaterite}), and then suspended in distilled water. A sample
of $200\ \mu L$ was pipptted and a single microsphere was captured via
a circular polarised beam.

Due to Van der Waal's forces some of the microspheres were stuck together, 
fortunately individual sphere's were still present. Multiple microspheres 
were trapped and their rotation rate was determined by looking at the 
peak frequency component of the collected power spectrum. The shear flow
was estimated using eq.\eqref{eq:birefringent_shear} assuming the spheres
were operating in the bulk fluid and away from any boundaries.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.82\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
	\end{subfigure}
	\begin{subfigure}{0.82\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
	\end{subfigure}
	\caption{(Top) Fluid flow radiating out from the surface of a rotating 
		vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
		is indicated by the dotted line. Vaterite radii and rotation frequencies 
		are shown, the laser power was kept constant at 450 mW. Reported rotation
		rates, and their corresponding fluid flow and shear rates, for vaterite 
		are also plotted alongside lab results. Results from \cite{Parkin2009, Arita2016} are included as well.}
	\label{fig:vaterite_shear}
\end{figure}

From Fig.\ref{fig:vaterite_shear} there is not a strong relationship between
particle size and rotation rate, this is contrary to much of the theoretical
predictions that predict an exponential decay with particle size. This can
be in part due to the fact that synthesising perfectly spherical spheres that
have uniform birefringence across the whole population is difficult. Despite 
our best efforts at controlling the growth rate the smallest particle ever 
synthesised was around $3\ \mu m$ in diameter. The vaterite spheres would 
often stick together while suspended in water after a short period of time. 
The fastest reported rotation rate found during this project was by 
\cite{Arita2016} that achieved a rotation rate of $5\ kHz$, this is plotted 
on Fig~\ref{fig:vaterite_shear} as the dotted line. Even at that extreme a 
rotation rate the region in which nucleation is at its optimal likelihood is 
only $20\ nm$ wide. This suggests that a more complicated fluid flow 
arrangement is required in order to achieve the necessary shear rate and over 
a larger volume of fluid.
